---
title: "Congaree Biosphere Reserve"
author: "Ethan Shafron"
date: "July 2, 2019"
output: 
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE, echo=FALSE, warning=FALSE, message=FALSE}

packages <- c("leaflet", "rgdal", "dplyr", "tidyr", "sf", "readxl", "here", "mapview", "raster")

if (length(setdiff(packages, rownames(installed.packages))) > 0){install.packages(setdiff(packages, rownames(installed.packages())))}

library(leaflet)
library(rgdal)
library(dplyr)
library(tidyr)
library(sf)
library(readxl)
library(here)

df <- file.info(list.files(here('data/input/updated'), full.names = TRUE))
best_file <- rownames(df)[which.max(df$mtime)]

good_marker_data <- st_read(best_file)
br <- st_read(here('data/input', 'AOI_BR_NAD.shp'))
br <- st_transform(br, 4326)


```



The Congaree Biosphere Reserve is not just unique for its natural landscape - it also has a deep cultural landscape like no other. From pre-colonial times, through slavery, the Civil War, and Reconstruction, to now, humans in this area have left indelible marks on the land that we now recognize as a part of the Biosphere Reserve. 


These historical markers remind us that the landscape does not contain one story of the land and one of the people, but rather a continuous narrative in which both the land and people play a part. We hope that this interactive tool highlights some of these place-based connections that various communities have here, and that it inspires you to create your own connections with the central midlands.


To use this map, simply zoom and pan around the area of interest. You can also click on "bunches" of markers to zoom in to a particular cluster of points. Hovering over a point will show you the name of the historical marker, and clicking it will reveal the text.


The data for this page came from [here](https://www.google.com/maps/d/viewer?mid=1t3m4iA_-a3hA8J_oPtxxNJJ2M_A&ll=33.69049557907872%2C-81.32880159895848&z=7). The Interactive map was developed using [R markdown](https://rmarkdown.rstudio.com/) and leaflet.





```{r, echo=FALSE, message = FALSE, warning=FALSE}
library(mapview)
library(raster)

leaflet(good_marker_data, width = 800, height = 800) %>%
  addProviderTiles(provider = providers$OpenStreetMap) %>%
  addMarkers(clusterOptions = markerClusterOptions(), popup = ~as.character(Text), label = ~as.character(`Marker.name...front`)) %>%
  addPolygons(data = br, fillOpacity = 0) %>%
  addHomeButton(ext = extent(br), layer.name = "Zoom to Biosphere Reserve", position = 'topright')

```


